---
title: "Portfolio"
author: "Faris Alsubaie"
date: "11/3/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
library(tidyverse)
library(ggplot2)
library(dplyr)
library(rsample)
library(recipes)
library(caret)
```

## Portfolio Builder Exercise #1
```{r}
#https://datahub.io/core/natural-gas#resource-natural-gas_zip
gas_prices <- read.csv("data/monthly_csv.csv")
gas_pricesD <- read.csv("data/daily_csv.csv")
```
```{r}
summary(gas_prices)
summary(gas_pricesD)
```
# Assess the distribution of the target / response variable.


```{r}

mean(gas_prices$Price)
ggplot(gas_pricesD, aes(x=Price))+
geom_histogram()
```

# Is the response skewed?

From the histogram above we can say it is positively skewed.

#Does applying a transformation normalize the distribution?
Yes, I used log10.

```{r}



Ngp_d <- log10(gas_pricesD$Price)

hist(log10(gas_pricesD$Price), breaks = 20)

summary(Ngp_d)

#ggplot(x = log10(gas_pricesD$Price))+
#geom_histogram()


#mod <- lm(Price ~ Date, data = gas_pricesD)
```
# Assess the dataset for missingness.
How many observations have missing values?
Plot the missing values. Does there appear to be any patterns to the missing values?
How do you think the different imputation approaches would impact modeling results?

 - their is 1 missing value
```{r cars}
sum(is.na(gas_pricesD))

gas_pricesD %>%
  is.na() %>%
  reshape2::melt() %>%
  ggplot(aes(Var2, Var1, fill=value)) + 
    geom_raster() + 
    coord_flip() +
    scale_y_continuous(NULL, expand = c(0, 0)) +
    scale_fill_grey(name = "", labels = c("Present", "Missing")) +
    xlab("Observation") +
    theme(axis.text.y  = element_text(size = 4))
```
```{r}
visdat::vis_miss(gas_pricesD, cluster = TRUE)
```


# Assess the variance across the features.
Do any features have zero variance?
Do any features have near-zero variance?

```{r}

```

